\bf Title

### 2023-10-26 article/paper_padic_spec.tex
variables en maj

\newcommand{\R}{\mathbb{R}} % real numbers
\newcommand{\C}{\mathbb{C}} % complex numbers
\newcommand{\N}{\mathbb{N}} % integers
\newcommand{\sym}{\mathbb{S}} % matrices symétriques
\renewcommand{\span}[1]{{\text{span}(#1)}} % simone's comments
\newcommand{\calL}{\mathcal{L}} % simone's comments

@@

@@

Let $\sym_d(\R)$ be the vector space of $d \times d$ real symmetric matrices. A matrix $M \in \sym_d(\R)$
is called {\it positive semidefinite} ($M \succeq 0$) whenever the quadratic
form $x \mapsto x^TMx$ is nonnegative for all $x\in \R^n$. By the Spectral Theorem, $M \succeq 0$
if and only if the eigenvalues of $M$ are nonnegative, and this is equivalent to all the principal minors
of $M$ being nonnegative. The set of positive semidefinite matrices, denoted $\sym_d^+(\R) \subset \sym_d(\R)$,
is a closed convex cone with non-empty interior in $\sym_m(\R)$.

Let $A_0,A_1,\ldots,A_n \in \sym_d(\R)$, and let $\calL = A_0+\span{A_1,\ldots,A_n}$ be the affine space
containing $A_0$ and with direction the vector space spanned by $A_1,\ldots,A_n$ (which we assume linearly
independent). The intersection
$\calL \cap \sym_d^+(\R) = \left\{A(x) := A_0+\sum_i x_i A_i \in \sym_d(\R) : A(x) \succeq 0\right\}$, or
equivalently, its pre-image $S = \{x \in \R^n : A(x) \succeq 0\}$ under the map $x \mapsto A(x)$, is called
a {\it real spectrahedron}. As affine sections of $\sym_d^+(\R)$, real spectrahedra are closed convex sets,
but might have empty interior in $\calL$ (resp. in $\R^n$). Moreover, they are basic semialgebraic sets,
defined by sign conditions on the principal minors of the defining matrix as well as sign conditions on the
coefficients of its characteristic polynomial (cf. \Cref{ell3}).

\begin{example}\label{ell3}
  Let $d=n=3$. The $3$-elliptope is the three-dimensional spectrahedron given by
  $$
  \mathcal{E}_3 :=
  \left\{
  x
  \in \R^3 :
  \begin{bmatrix}
    1 & x_1 & x_2 \\
    x_1 & 1 & x_3 \\
    x_2 & x_3 & 1
  \end{bmatrix}
  \succeq 0
  \right\}
  =
  \left\{
  x
  \in \R^3 :
  \begin{array}{r}
    3-x_1^2-x_2^2-x_3^2 \geq 0 \\
    1+2x_1x_2x_3-x_1^2-x_2^2-x_3^2 \geq 0
  \end{array}
  \right\}.
  $$
  The polynomials on the right are the (non-constant) coefficients of the univariate polynomial
  $p(t) = \det(A(x)+t I)$,
  whose roots are the opposites of the eigenvalues of the defining matrix. Equivalently $\mathcal{E}_3$
  is defined by positivity of the principal minors of the defining matrix:
  $$
  1-x_1^2 \geq 0, \,\,\, 1-x_2^2 \geq 0, \,\,\, 1-x_3^2 \geq 0 \,\,\, \text{ and } \,\,\, 1+2x_1x_2x_3-x_1^2-x_2^2-x_3^2 \geq 0
  $$
  \exend
\end{example}

The optimization problem of miminizing a linear function over a real spectrahedron is called {\it semidefinite programming} (SDP). Given $C,A_0,A_1,\ldots,A_n \in \sym_m(\R)$, the corresponding semidefinite program is defined as
\begin{equation}
  \label{SDP}
\begin{array}{rcll}
  p^* & := & \inf_{x \in \R^n} & \left\langle c, x \right\rangle \\
  &    & \text{s.t.}         & A_0+\sum_{i=1}^n x_i A_i \succeq 0.
\end{array}
\end{equation}
Real polyhedra (subsets of $\R^n$ defined by finitely many affine inequalities) are special cases of real
spectrahedra, where the defining matrices $A_0, \ldots, A_n$ can be chosen diagonal. Thus SDP is a generalization
of linear programming (LP). In a fixed-precision model of computation, SDP is solvable in polynomial time in the
input size (matrix size,
number of variables, bit size of the coefficients), in $\log(\frac{1}{\varepsilon})$ (where $\varepsilon$ is the
precision) and $\log(\frac{1}{R})$, where $R$ is an {\it a priori} upper bound on the norm of the solution,
cf. \cite[Sec.1.9]{deKlerk}. In exact arithmetic, the complexity status of SDP is essentially open but complexity
bounds are known based either on general quantifier elimination \cite{ramana1997exact,porkolab1997complexity}
or by exploiting the determinantal structure of the constraints \cite{henrion2016exact}.

A more general class of convex semialgebraic sets are obtained by linear projections of real spectrahedra
called {\it semidefinite representable sets} (or spectrahedral shadows): these are sets of the form
$$
P = \left\{x = \left[\begin{smallmatrix} x_1 \\ \vdots \\ x_n \end{smallmatrix}\right] \in \R^n : \exists\,y\in\R^m, \, A_0 + \sum_{i=1}^n x_i A_i + \sum_{j=1}^m y_j B_j \succeq 0\right\}
$$

\begin{example}
\label{fermat_quartic}
The basic closed semialgebraic set $P = \{(x_1,x_2) \in \R^2 : 1-x_1^4-x_2^4 \geq 0\}$ is not a spectrahedron
but it is semidefinite representable:
$$
P = \left\{x = \begin{bmatrix} x_1 \\ x_2 \end{bmatrix} \in \R^2 :
\exists\, (y_1,y_2) \in \R^2, \,
\begin{bmatrix}
  1+y_1 & y_2 \\
  y_2 & 1-y_1
\end{bmatrix}
\bigoplus
\begin{bmatrix}
  1 & x_1 \\
  x_1 & y_1
\end{bmatrix}
\bigoplus
\begin{bmatrix}
  1 & x_2 \\
  x_2 & y_2
\end{bmatrix}
\succeq 0
\right\}
$$
\end{example}

@@

\begin{definition}
  Let $\ell_1,\ldots,\ell_d \in K[X]_1$ be linear inequalities on $X = (X_1, \ldots, X_n)$.
  A {\it polyhedron} is a subset $P \subset K^s$ defined by linear inequalities:
  $P = \val\left( \ell_1(x)\right)\ge 0$,\ldots, $ \val\left( \ell_n\left( x \right) \right)\ge 0$ for $ \ell_1,\ldots, \ell_n $ fixed linear forms and for all $x \in K ^s$.
\end{definition}

\begin{example}
	The closed unit ball of $K^s$ for the infinity norm [not sure if it is really used ] is convex polyhedron.It is the set of $\left( x_1,\ldots,x_s \right) \in  ^s$ so that $$
	\begin{pmatrix} x_1 &  & & \\
		& x_2 &0 &  \\
		&  0& \ddots & \\
		&  & & x_s/
	\end{pmatrix} \ge 0.$$ which is in fact $\OK^s$.

\end{example}

%\subsection{linear programming}
	We call { \it $p$-adic linear programming} the problem :
\begin{equation}
	\tag{PLp}
	\begin{matrix}
		\text{Minimize } \val\left(\left<c,x \right>\right) \text{ subject to }\\
		Ax + b \ge 0
	\end{matrix}
	\label{eqn:Proglinp}
\end{equation}

@@

\subsection{Definition}
\newcommand\mat{postive semidefinite matrix } 
\newcommand\Mat{Positive semidefinite matrix }
\newcommand\mats{positive semidefinite matrices }
\newcommand\Mats{positive semidefinite matrices }
\begin{definition}
	We call {\it \mat} any matrix $M \in \mathcal{M}_n\left( K \right) $ whose eigenvalues have nonegative valuation (in $\overline{K}$) .

>>>>

\newcommand{\R}{\mathbb{R}} % real numbers
\newcommand{\C}{\mathbb{C}} % complex numbers
\newcommand{\N}{\mathbb{N}} % integers
\newcommand{\allmat}{\mathbb{M}} % matrices
\newcommand{\sym}{\mathbb{S}} % matrices symétriques
\renewcommand{\span}[1]{{\text{span}(#1)}} % simone's comments
\newcommand{\calL}{\mathcal{L}} % simone's comments

@@

The $K$-vector space of polynomials of degree $\leq d$ on $X=(X_1,\ldots,X_n)$
with coefficients in $K$ is denoted by $K[X]_{d}$. The vector space of matrices of
size $p \times q$ with entries in $K$ is denoted by $\allmat_{p,q}(K)$. For the sake
of ease of writing, diagonal matrices with entries $a_1,\ldots,a_d$ are denoted by
$a_1 \oplus \cdots \oplus a_n$.

@@

Let $\sym_d(\R) \subset \allmat_{d,d}(\R)$ be the vector space of $d \times d$ real symmetric
matrices. A matrix $M \in \sym_d(\R)$
is called {\it positive semidefinite} ($M \succeq 0$) whenever the quadratic
form $x \mapsto x^TMx$ is nonnegative for all $x\in \R^n$. By the Spectral Theorem, $M \succeq 0$
if and only if the eigenvalues of $M$ are nonnegative, and this is equivalent to all the principal minors
of $M$ being nonnegative. The set of positive semidefinite matrices, denoted $\sym_d^+(\R) \subset \sym_d(\R)$,
is a closed convex cone with non-empty interior in $\sym_d(\R)$.

Let $A_0,A_1,\ldots,A_n \in \sym_d(\R)$, and let $\calL = A_0+\span{A_1,\ldots,A_n}$ be the affine space
containing $A_0$ and with direction the vector space spanned by $A_1,\ldots,A_n$ (which we assume linearly
independent). The intersection
$\calL \cap \sym_d^+(\R) = \left\{A(X) := A_0+\sum_i X_i A_i \in \sym_d(\R) : A(X) \succeq 0\right\}$, or
equivalently, its pre-image $S = \{X \in \R^n : A(X) \succeq 0\}$ under the map $X \mapsto A(X)$, is called
a {\it real spectrahedron}. As affine sections of $\sym_d^+(\R)$, real spectrahedra are closed convex sets,
but might have empty interior in $\calL$ (resp. in $\R^n$). Moreover, they are basic semialgebraic sets,
defined by sign conditions on the principal minors of the defining matrix as well as sign conditions on the
coefficients of its characteristic polynomial (cf. \Cref{ell3}).

\begin{example}\label{ell3}
  Let $d=n=3$. The $3$-elliptope is the three-dimensional spectrahedron given by
  $$
  \mathcal{E}_3 :=
  \left\{
  X
  \in \R^3 :
  \begin{bmatrix}
    1 & X_1 & X_2 \\
    X_1 & 1 & X_3 \\
    X_2 & X_3 & 1
  \end{bmatrix}
  \succeq 0
  \right\}
  =
  \left\{
  X
  \in \R^3 :
  \begin{array}{r}
    3-X_1^2-X_2^2-X_3^2 \geq 0 \\
    1+2X_1X_2X_3-X_1^2-X_2^2-X_3^2 \geq 0
  \end{array}
  \right\}.
  $$
  The polynomials on the right are the (non-constant) coefficients of the univariate polynomial
  $p(T) = \det(A(X)+T I_3)$,
  whose roots are the opposites of the eigenvalues of the defining matrix. Equivalently $\mathcal{E}_3$
  is defined by positivity of the principal minors of the defining matrix:
  $$
  1-X_1^2 \geq 0, \,\,\, 1-X_2^2 \geq 0, \,\,\, 1-X_3^2 \geq 0 \,\,\, \text{ and } \,\,\, 1+2X_1X_2X_3-X_1^2-X_2^2-X_3^2 \geq 0
  $$
  \exend
\end{example}

The optimization problem of miminizing a linear function over a real spectrahedron is called {\it semidefinite programming} (SDP). Given $c \in \R^n$ and $A_0,A_1,\ldots,A_n \in \sym_d(\R)$, the corresponding semidefinite program is defined (in standard dual form) as
\begin{equation}
  \label{SDP}
\begin{array}{rcll}
  p^* & := & \inf_{X \in \R^n} & \left\langle c, X \right\rangle \\
  &    & \text{s.t.}         & A_0+\sum_{i=1}^n X_i A_i \succeq 0.
\end{array}
\end{equation}
Real polyhedra (subsets of $\R^n$ defined by finitely many affine inequalities) are special cases of real
spectrahedra, where the defining matrices $A_0, \ldots, A_n$ can be chosen diagonal, so that $A(X) =
\ell_1(X) \oplus \cdots \oplus \ell_d(X)$.

In other words, SDP is a generalization
of Linear Programming (LP). In a fixed-precision model of computation, SDP is solvable in polynomial time in the
input size (matrix size,
number of variables, bit size of the coefficients), in $\log(\frac{1}{\varepsilon})$ (where $\varepsilon$ is the
precision) and $\log(\frac{1}{R})$, where $R$ is an {\it a priori} upper bound on the norm of the solution,
cf. \cite[Sec.1.9]{deKlerk}. In exact arithmetic, the complexity status of SDP is essentially open but complexity
bounds are known based either on general quantifier elimination \cite{ramana1997exact,porkolab1997complexity}
or by exploiting the determinantal structure of the constraints \cite{henrion2016exact}.

A more general class of convex semialgebraic sets are obtained by linear projections of real spectrahedra
called {\it semidefinite representable sets} (or spectrahedral shadows): these are sets of the form
$$
P = \left\{X = \left[\begin{smallmatrix} X_1 \\ \vdots \\ X_n \end{smallmatrix}\right] \in \R^n : \exists\,Y\in\R^p, \, A_0 + \sum_{i=1}^n X_i A_i + \sum_{j=1}^p Y_j B_j \succeq 0\right\}
$$

\begin{example}
\label{fermat_quartic}
The basic closed semialgebraic set $P = \{(X_1,X_2) \in \R^2 : 1-X_1^4-X_2^4 \geq 0\}$ is not a spectrahedron
but it is semidefinite representable:
$$
P = \left\{X = \begin{bmatrix} X_1 \\ X_2 \end{bmatrix} \in \R^2 :
\exists\, (Y_1,Y_2) \in \R^2, \,
\begin{bmatrix}
  1+Y_1 & Y_2 \\
  Y_2 & 1-Y_1
\end{bmatrix}
\bigoplus
\begin{bmatrix}
  1 & X_1 \\
  X_1 & Y_1
\end{bmatrix}
\bigoplus
\begin{bmatrix}
  1 & X_2 \\
  X_2 & Y_2
\end{bmatrix}
\succeq 0
\right\}
$$
\end{example}

@@

\begin{definition}
  A {\it polyhedron} is a subset $P \subset K^n$ of the form
  $P = \{X \in K^n : \val(\ell_1(X))\geq0,\ldots,\val(\ell_n(X))\geq 0\}$,
  where $\ell_1,\ldots,\ell_d \in K[X]_1$ are affine polynomials on $X = (X_1, \ldots, X_n)$.
\end{definition}

\begin{example}
  The closed unit ball in $K^n$ for the supremum norm $||\cdot||_{\infty}$ is the polyhedron defined as
  the set of $(X_1,\ldots,X_n) \in K^n$ such that
  $$
  \begin{pmatrix} x_1 &  & & \\
    & x_2 &0 &  \\
    &  0& \ddots & \\
    &  & & x_s/
  \end{pmatrix} \ge 0.$$ which is in fact $\OK^s$.
\end{example}

%\subsection{linear programming}

A {\it linear programming problem} (or simply {\it linear program}) is
\begin{equation}
  \tag{LP}
  \begin{matrix}
    \text{Minimize } \val\left(\left<c,x \right>\right) \text{ subject to }\\
    Ax + b \ge 0
  \end{matrix}
  \label{eqn:Proglinp}
\end{equation}

@@

\subsection{Definition}
%\newcommand\mat{postive semidefinite matrix } 
\newcommand\Mat{Positive semidefinite matrix }
\newcommand\mats{positive semidefinite matrices }
\newcommand\Mats{positive semidefinite matrices }
\begin{definition}
	We call ... any matrix $M \in \mathcal{M}_n\left( K \right) $ whose eigenvalues have nonegative valuation (in $\overline{K}$) .
